---
title: "Assignment1"
author: "Philip & Philipp"
date: "23 Feb 2016"
output: pdf_document
bibliography: Ass1Packages.bib
---

```{r, prompt=FALSE}
# Load packages and create BibTeX file for R-packages
PackagesUsed <- c("ggplot2", "repmis", "doBy")

setwd("~/Documents/Collaborative Social Science/Collaborative Analysis Assignments/PandP_Ass1")

# Load PackagesUsed and create .bib BibTeX file
repmis::LoadandCite(PackagesUsed, file = "Ass1Packages.bib", install = FALSE)
```

# Analysis of the dataset 'occupationalStatus'
The dataset consists of a contingency table between the occupational status measured on an 8-point scales for fathers and their sons. 

```{r, include=FALSE}
source('Assignment1_rscript.R')
```


``` {r}
dist_combined <- rbind(dist_fathers, dist_sons)
barplot(dist_combined, 
        col=c("navyblue", "darkkhaki"),
        main="Distribution of occupational status among fathers and sons",
        xlab = "Occupational status categories",
        legend = c("Fathers", "Sons"),
        ylim = c(0, 1600),
        beside = TRUE
        )
```

Figure 1 shows the distribution of occupational status for sons and fathers respectively. There does not appear to be any major generational shifts, though sons are slightly overrepresented in occupational status group 7 and 8. 

```{r}
knitr::kable(frequency_table, digits = 2)
```

Table 1 is equivalent to figure 1 except that it shows the frequency distribution for each status category. 

## Generational mobility
Figure 1 and table 1 show the overall distribution of occupational status, but is not informative on the extent to which there are generational mobility. Figure 2 shows the distribution of generational mobility. -7 indicates observations where the father had social status 8 and the son 1. As such, a zero is when father and son had the same occupational status. 

```{r}
barplot(collapsed$Freq.sum, 
        names = collapsed$difference,
        col = "Navyblue",
        ylim = c(0, 1400)
        )
```

Packages used [@CiteR], [@R-doBy], [@R-ggplot2] and [@R-repmis].

## References
